### 2016-12-29 paper/paris17/paper.tex
add SConstruct

Explicit finite difference modeling is constrained by figuring  out a sampling in time and space that results in stable propagation and does not create dispersive events.  For stability the Courant-Friedrichs-Lewy condition \cite[]{courant1967partial} must be met.  Stability is a function of limiting what percentage of a grid cell energy can move in one time step. Stability is therefore a function of the maximum velocity $v_{max}$, the minimum spatial sampling $d_{min}$, and the time step $dt$. For stability, we have $v_{max}\frac{dt}{d_{min}} < .5$.

The stability condition pushes one to use larger spatial sampling (faster, but less resolution) and/or finer time sampling (more expensive).  Dispersion, on the other hand, is a function of the minimum velocity $v_{min}$, the maximum frequency $f_{max}$, and the maximum spaital sampling $d_{max}$.   To avoid grid dispersion we need to  sample a given frequency with a minimum number of points.  There isn't a consensus on the minimum number
of points. For the purpose of this paper I will require  3.2  points therefore, we have $\frac{v_{min} }{f_{max}d_{max}} > 3.2 \label{eq:dispersion}$.

@@

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[height=1.3in]{./fig/domain-decompose.pdf}
        \caption{Domain decomposition}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[height=1.3in]{./fig/inner-outer.pdf}
        \caption{Task decomposition}
    \end{subfigure}
    ~ 
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[height=1.1in]{./fig/overlap.pdf}
        \caption{Comp. \& Comm. overlapping}
    \end{subfigure}
    \caption{Computation and communication overlapping over multiple GPUs.}
\end{figure}

@@

Figure xxx shows the wave-field using the standard acoustic wave equation (left) and an attenuation wave-field (right). Note the difference in the frequency content. This is can be more clearly seen in Figure xxx which shows the spectrum of wave-field at 0.9 seconds with $Q_p = Q_s = 20$. Notice the energy decay of frequency over time. The key observation is that there is no reason to worry about dispersion at frequencies that have attenuated. Specifically, I redo my dispersion calculation (equation 2) several times while propagating a wave-field. At each time block I use a fmax based on frequencies whose energy has not been reduced more than some percentage (in this paper 96\%). As a result, as I forward propagate in time my grid cells, and if I desire, time sampling, get larger. Figure 3 shows the speedup factor (defined as the number of grid cells times time steps) as a function of propagation time for different Q values. The longer the time record, the more using Q pays off in terms speed up. Using this approach the early time steps dominate the computation. For example assuming an initial maximum frequency of 90Hz and a constant velocity of 2000m/s, a Q value of 200, the total speedup is 4 even though most of the propagation time shows significantly more speedup. Figure 4 shows the speedups for Q values ranging from 150 to 350.

>>>>

Explicit finite difference modeling is constrained by figuring  out a sampling in time and space that results in stable propagation and does not create dispersive events.  For stability the Courant-Friedrichs-Lewy condition \cite[]{courant1967partial} must be met.  Stability is a function of limiting what percentage of a grid cell energy can move in one time step. Stability is therefore a function of the maximum velocity $v_{max}$, the minimum spatial sampling $d_{min}$, and the time step $dt$. For stability, we have $v_{max}\frac{dt}{d_{min}} < .5$.  The stability condition pushes one to use larger spatial sampling (faster, but less resolution) and/or finer time sampling (more expensive).  Dispersion, on the other hand, is a function of the minimum velocity $v_{min}$, the maximum frequency $f_{max}$, and the maximum spaital sampling $d_{max}$.   To avoid grid dispersion we need to  sample a given frequency with a minimum number of points.  There isn't a consensus on the minimum number
of points. For the purpose of this paper I will require  3.2  points therefore, we have $\frac{v_{min} }{f_{max}d_{max}} > 3.2 \label{eq:dispersion}$.

@@

Thirdly, we design a computation and communication overlapping scheme to maximize the concurrency of different resources. The entire wavefield domain is divided into 4 subdomain, each assigning to a GPU shown in Figure \ref{fig:domain-decomposition}. Note that we need ghost cells to store and exchanges the halos, denoted by the arrows. In each subdomain, we further divide it into three parts, the halo part, the outer part and the inner part shown in Figure \ref{fig:task-decomposition}. The computation of inner and outer parts are assigned to different CUDA streams so that they are executed concurrently. When the computation of the outer part is finished, the GPU starts exchanging halos with its neighbors. In this way, the computation of the inner part can hide the communication shown in Figure \ref{fig:overlap}.

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[height=1.3in]{./fig/domain-decompose.pdf}
        \caption{Domain decomposition}
        \label{fig:domain-decomposition}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[height=1.3in]{./fig/inner-outer.pdf}
        \caption{Task decomposition}
        \label{fig:task-decomposition}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[height=1.1in]{./fig/overlap.pdf}
        \caption{Comp. \& Comm. overlapping}
        \label{fig:overlap}
    \end{subfigure}
    \caption{Computation and communication overlapping over multiple GPUs.}
\end{figure}

@@

Figure xxx shows the wave-field using the standard acoustic wave equation (left) and an attenuation wave-field (right). Note the difference in the frequency content. This is can be more clearly seen in Figure xxx which shows the spectrum of wave-field at 0.9 seconds with $Q_p = Q_s = 20$.

Notice the energy decay of frequency over time. The key observation is that there is no reason to worry about dispersion at frequencies that have attenuated. Specifically, I redo my dispersion calculation (equation 2) several times while propagating a wave-field. At each time block I use a fmax based on frequencies whose energy has not been reduced more than some percentage (in this paper 96\%). As a result, as I forward propagate in time my grid cells, and if I desire, time sampling, get larger. Figure 3 shows the speedup factor (defined as the number of grid cells times time steps) as a function of propagation time for different Q values. The longer the time record, the more using Q pays off in terms speed up. Using this approach the early time steps dominate the computation. For example assuming an initial maximum frequency of 90Hz and a constant velocity of 2000m/s, a Q value of 200, the total speedup is 4 even though most of the propagation time shows significantly more speedup. Figure 4 shows the speedups for Q values ranging from 150 to 350.
